---
title: "Week 4 Lab"
author: "Shannon Duffy"
date: "9/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Methicillin-resistant Staphylococcus aureus or MRSA can cause life-threatening infections. Suppose that the death rate from MRSA is 20%. Twenty patients at a hospital are diagnosed with MRSA.
Find the following probabilities:
i. Less than 4 in this group will die
```{r}
pbinom(3, 20, 0.2)
```
ii. Between 3 and 6 (inclusive of 3 and 6) will die
```{r}
sum(dbinom(3:6, 20, 0.2))
```
iii. More than 1 will die
```{r}
1-pbinom(1,20,0.2)
```

2. The V1 Flying Bomb (aka Buzzbomb) was a precursor of the modern cruise missile and was used by Germany in WWII to bomb
England. Historians have analyzed the intensity of bombing in
England and found that each 0.25-km2 area in London received an average of 0.929 bomb strikes during the war. Assuming that the bombing was done randomly over space, how many of the 550 0.25-km2regions of England would you expect to have had been hit by two or more bombs?
```{r}
(1-ppois(1,0.929))*550
```

3. Assume that the mean number of frog calls heard per hour when conducting a swamp survey is 2.2. If frogs call randomly through the night, what is the chance that a certain hour has no calls at all? Now, determine the chance that all six hours of a typical 6-hour survey would be completely silent (thus boring you out of your skull) assuming that frog calling is independent from one hour to the next.
```{r}
(dpois(0,2.2))^6
```

4. You read that the rate of albinism in wild hogs is 20%. How unusual would it be to find 10 or more albino hogs in a sample 25? Express as a percentage.
```{r}
1-pbinom(9, 25, .20)
```

5. Imagine that the probability of infection with rust fungus is 40% in a plant population. You go out and sample five plants at a time (because this is how many fit into your bucket) to look for infected individuals.
i. Make a nice graph to depict the probabilities of having 0, 1,…, 5 infected individuals in a sample.
```{r}
barplot(dbinom(0:5,5, 0.4), names.arg = c(0:5),
        xlab = "Plants with fungus",
        ylab = "Probability")
abline(h=0)
```
ii. What is the probability of obtaining 3 or more infected individuals in a group of 5?
```{r}
1-pbinom(2, 5, 0.4)
```
problems for lab 4 2
6. The heart rate of the European common shrew, Sorex araneus, has been clocked at an average rate 528 beats per minute with a standard deviation of 55 beats per minute. Data from a wild population suggested a normal distribution for this random variable. You measure the heart rate of an American common shrew, Sorex cinereus at 480 BPM. How unlikely would this be if the distribution of heart rates for S. cinereus was the same as for S. araneus?
i. How unusually low would that be?
```{r}
pnorm(480, 528, 55)
```
ii. How unusual period (high or low) would that be? (That is, how unusually far from the mean is this?)
```{r}
pnorm(480, 528, 55)*2
(1-pnorm(576, 528, 55)) + pnorm(480, 528, 55)
```

7. Last week you calculated the mass of a bunch of least shrews at 5.759 g with a standard deviation of 0.848 g. Assume that the distribution of mass in this species is normal.
i. What is the chance of observing a shrew in this population with a mass above 6 g?
```{r}
1-pnorm(6, 5.759, 0.848)
```
ii. What is the chance of finding a shrew weighing less than 4 g?
```{r}
pnorm(4, 5.759, 0.848)
```
iii. What is the chance of finding a shrew weighing between 5 and 6 g?
```{r}
pnorm(6, 5.759, 0.848)-pnorm(5, 5.759, 0.848)
```
iv. What is the weight that delimits the upper 2.5% of the distribution of masses? That is, what is the weigh above which you are only 2.5% likely to find any shrews?
```{r}
qnorm(0.975, 5.759, 0.848)
```
